### 2024-02-15 paperAndPoster/poster.tex
trouble with bibliography

\documentclass[landscape]{tikzposter} % See Section 3
\input{common.tex}

\title{\actualTitle} % See Section 4.1
\author{Eric Campbell \\ \texttt{eric.robert.campbell@gmail.com}} 
\usetheme{Basic} % See Section 5

\begin{document}
\maketitle % See Section 4.1
\block{Abstract}{Put abstract here} % See Section 4.2
\begin{columns} % See Section 4.4
\column{0.333333333} % See Section 4.4
\block{BlocktitleB}{Blocktext}

\column{0.333333333}
\block{BlocktitleC}{Blocktext}

\column{0.333333333}
\block{Results}{Put Results Here}
\block{Next Steps}{Ideas for next steps here}
\block{Bibliography}{
    % \input{paper.bbl}

    For all data and code used here, along with additional analyses, see \url{https://github.com/EricRobertCampbell/aps-2024-poster-sexual-dimorphism}.
}
\note{Notetext} % See Section 4.3

\end{columns}
\end{document}