### 2024-02-16 paperAndPoster/poster.tex
feat: inserted paper bibliography into poster

\documentclass[landscape]{tikzposter} % See Section 3
\usepackage[style=authoryear-comp, locallabelwidth]{biblatex}
\usepackage{anyfontsize}
\input{common.tex}

\renewcommand*{\bibfont}{\tiny}

\title{\actualTitle} % See Section 4.1
\author{Eric Campbell \\ \texttt{eric.robert.campbell@gmail.com}} 
\usetheme{Basic} % See Section 5

\begin{document}
\maketitle % See Section 4.1
\block{Abstract}{Put abstract here} % See Section 4.2
\begin{columns} % See Section 4.4
\column{0.333333333} % See Section 4.4
\block{BlocktitleB}{Blocktext}

\column{0.333333333}
\block{BlocktitleC}{Blocktext}

\column{0.333333333}
\block{Results}{Put Results Here}
\block{Next Steps}{Ideas for next steps here}
\block{Bibliography}{
    \makeatletter
    \def\blx@bblfile@biber{%
      \blx@secinit
      \begingroup
      \blx@bblstart
        \input{poster.bbl}
      \blx@bblend
      \endgroup
      \csnumgdef{blx@labelnumber@\the\c@refsection}{0}}
    \makeatother
    \begin{minipage}{\linewidth}
        \begin{center}\mbox{}\vspace{0pt}
            \printbibliography[heading=none]
        \end{center}
        \vspace{\baselineskip}
    \end{minipage}

    For all data and code used here, along with additional analyses, see \url{https://github.com/EricRobertCampbell/aps-2024-poster-sexual-dimorphism}.
}
\note{Notetext} % See Section 4.3

\end{columns}
\end{document}